---
title: "Diamond sizes"
date: 2016-08-25
output: html_document
---

```{r setup, include = FALSE}
library(ggplot2)
library(dplyr)

smaller <- diamonds %>% 
  filter(carat <= 2.5)
```

We have data about `r nrow(diamonds)` diamonds. Only 
`r nrow(diamonds) - nrow(smaller)` are larger than
2.5 carats. The distribution of the remainder is shown
below:

```{r, echo = FALSE}
smaller %>% 
  ggplot(aes(carat)) + 
  geom_freqpoly(binwidth = 0.01)
```

---
27.3.1 Exercise 3
---
title: "Penny Wang"
output: html_document
---
## Education
* **2027**(Expected): B.S. in Computer Science (Mathematics minor) CUNY Queens College

## Experience
* **2026**: Microsoft Research Data Science Summer School (DS3) 
* **2026**: Breakthrough Tech AI/ML Fellow Cornell Tech 

---
Footnote, ^[I am a footnote.]. 
Footnote meaning according to the dictionary, ^[“A note of reference, explanation, or comment usually placed below the text on a printed page.” - Merriam Webster ]

> I am a block quote. 
>”Block quotation is a method of formatting to highlight sections of directly quoted text in your writing.” - University of Illinois Springfield

---

27.4.7 Exercise 1
---
title: "Diamond Sizes Analysis" 
output: html_document 
---
```{r setup-global}
library(ggplot2)
library(dplyr)
library(scales) 

knitr::opts_chunk$set(
  echo = FALSE
)

smaller <- diamonds %>% filter(carat <= 2.5) 

total_diamonds <- nrow(diamonds) 
larger_diamonds <- total_diamonds - nrow(smaller) 
percent_larger <- (larger_diamonds / total_diamonds) * 100 

top_20_diamonds <- diamonds %>%
  arrange(desc(carat)) %>%
  slice(1:20) %>%
  select(carat, cut, color, clarity, depth, table, price)
  ```

```{r smaller_plot}
smaller %>% 
  ggplot(aes(x = carat)) + 
  geom_histogram(binwidth = 0.01, fill = "steelblue", color = "white")
```

```{r top_20_diamonds_table}
knitr::kable(
  top_20_diamonds, 
  caption = "Top 20 Largest Diamonds by Carat Weight and Quality Metrics",
  col.names = c("Carat Weight", "Cut Quality", "Color Grade", "Clarity Rating", "Depth %", "Table %", "Price ($)")
)
```

```{r chunk_a, cache = TRUE} 
# Base Case
a_time <- lubridate::now()
print(paste("Chunk A executed at:", a_time))
```

```{r chunk_b, cache = TRUE, dependson = "chunk_a"} 
# Needs Parent A
b_time <- lubridate::now()
print(paste("Chunk B executed at:", b_time))
```

```{r chunk_c, cache = TRUE, dependson = "chunk_a"} 
# Needs Parent A
c_time <- lubridate::now()
print(paste("Chunk C executed at:", c_time))
```

```{r chunk_d, cache = TRUE, dependson = c("chunk_b", "chunk_c")} 
# Needs Both Parents B and C
d_time <- lubridate::now()
print(paste("Chunk D executed at:", d_time))
```
---