---
title: "3.6.2 Simple Linear Regression"
output: html_document
---

The ISLR2 library contains the Boston data set, which records medv 
(median house value) for 506 census tracts in Boston. We will seek 
to predict medv using 12 predictors such as rm (average number of rooms 
per house), age (proportion of owner-occupied units built prior to 1940) 
and lstat (percent of households with low socioeconomic status).

```{r}
library(ISLR2)
head(Boston)
```

To find out more about the data set, we can type `?Boston`.

We will start by using the `lm()` function to fit a simple linear 
regression model, with medv as the response and lstat as the predictor. 
The basic syntax is `lm(y ~ x, data)`, where y is the response, x is 
the predictor, and data is the data set in which these two variables are kept.

```{r}
lm.fit <- lm(medv ~ lstat, data = Boston)
attach(Boston)
lm.fit <- lm(medv ~ lstat)
```

If we type `lm.fit`, some basic information about the model is output. 
For more detailed information, we use `summary(lm.fit)`. This gives us 
p-values and standard errors for the coefficients, as well as the R2 
statistic and F-statistic for the model.

```{r}
lm.fit
summary(lm.fit)
```

We can use the `names()` function in order to find out what other pieces of 
information are stored in `lm.fit`. Although we can extract these quantities 
by name---e.g. `lm.fit$coefficients`---it is safer to use the extractor 
functions like `coef()` to access them.

```{r}
names(lm.fit)
coef(lm.fit)
```

In order to obtain a confidence interval for the coefficient estimates, 
we can use the `confint()` command.

```{r}
confint(lm.fit)
```

The `predict()` function can be used to produce confidence intervals and 
prediction intervals for the prediction of medv for a given value of lstat.

```{r}
predict(lm.fit, data.frame(lstat = (c(5, 10, 15))),
        interval = "confidence")

predict(lm.fit, data.frame(lstat = (c(5, 10, 15))),
        interval = "prediction")
```

For instance, the 95% confidence interval associated with a lstat value of 10 
is (24.47, 25.63), and the 95% prediction interval is (12.828, 37.28). As expected, 
the confidence and prediction intervals are centered around the same point 
(a predicted value of 25.05 for medv when lstat equals 10), but the latter are substantially wider.

We will now plot medv and lstat along with the least squares regression line 
using the `plot()` and `abline()` functions.

```{r}
plot(lstat, medv)
abline(lm.fit)
```

There is some evidence for non-linearity in the relationship between lstat and medv. 
We will explore this issue later in this lab.

The `abline()` function can be used to draw any line, not just the least squares 
regression line. To draw a line with intercept a and slope b, we type `abline(a, b)`. 
Below we experiment with some additional settings for plotting lines and points. 
The `lwd = 3` command causes the width of the regression line to be increased by a 
factor of 3; this works for the `plot()` and `lines()` functions also. We can also 
use the `pch` option to create different plotting symbols.

```{r}
abline(lm.fit, lwd = 3)
abline(lm.fit, lwd = 3, col = "red")
plot(lstat, medv, col = "red")
plot(lstat, medv, pch = 20)
plot(lstat, medv, pch = "+")
plot(1:20, 1:20, pch = 1:20)
```

Next we examine some diagnostic plots, several of which were discussed in Section 3.3.3. 
Four diagnostic plots are automatically produced by applying the `plot()` function 
directly to the output from `lm()`. In general, this command will produce one plot 
at a time, and hitting Enter will generate the next plot. However, it is often convenient 
to view all four plots together. We can achieve this by using the `par()` and `mfrow()` 
functions, which tell R to split the display screen into separate panels so that multiple 
plots can be viewed simultaneously. For example, `par(mfrow = c(2, 2))` divides the plotting 
region into a 2 x 2 grid of panels.

```{r}
par(mfrow = c(2, 2))
plot(lm.fit)
```

Alternatively, we can compute the residuals from a linear regression fit using the `residuals()` 
function. The function `rstudent()` will return the studentized residuals, and we can use this 
function to plot the residuals against the fitted values.

```{r}
plot(predict(lm.fit), residuals(lm.fit))
plot(predict(lm.fit), rstudent(lm.fit))
```

On the basis of the residual plots, there is some evidence of non-linearity. Leverage statistics 
can be computed for any number of predictors using the `hatvalues()` function.

```{r}
plot(hatvalues(lm.fit))
which.max(hatvalues(lm.fit))
```

The `which.max()` function identifies the index of the largest element of a vector. In this case, 
it tells us which observation has the largest leverage statistic.
