---
title: "ISLR2 Lab 5.3.1–5.3.3: Cross-Validation"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
install.packages("ISLR2")
library(ISLR2)
install.packages("boot")
library(boot)
```

## 5.3.1 The Validation Set Approach

We explore the use of the validation set approach in order to estimate the
test error rates that result from fitting various linear models on the Auto
data set.

Before we begin, we use the set.seed() function in order to set a seed for seed
R’s random number generator, so that the reader of this book will obtain
precisely the same results as those shown below. It is generally a good idea
to set a random seed when performing an analysis such as cross-validation
that contains an element of randomness, so that the results obtained can
be reproduced precisely at a later time.

We begin by using the sample() function to split the set of observations 
into two halves, by selecting a random subset of 196 observations out of
the original 392 observations. We refer to these observations as the training
set.

```{r}
set.seed(1)
train <- sample(392, 196)
```

(Here we use a shortcut in the sample command; see ?sample for details.)
We then use the subset option in lm() to fit a linear regression using only
the observations corresponding to the training set.

```{r}
lm.fit <- lm(mpg ~ horsepower, data = Auto, subset = train)
```

We now use the predict() function to estimate the response for all 392
observations, and we use the mean() function to calculate the MSE of the
196 observations in the validation set. Note that the-train index below
selects only the observations that are not in the training set.

```{r}
attach(Auto)
mean((mpg - predict(lm.fit, Auto))[-train]^2)
```

Therefore, the estimated test MSE for the linear regression fit is 23.27.We
can use the poly() function to estimate the test error for the quadratic and
cubic regressions.
```{r}
lm.fit2 <- lm(mpg ~ poly(horsepower, 2), data = Auto, subset = train)
mean((mpg - predict(lm.fit2, Auto))[-train]^2)

lm.fit3 <- lm(mpg ~ poly(horsepower, 3), data = Auto, subset = train)
mean((mpg - predict(lm.fit3, Auto))[-train]^2)
```

These error rates are 18.72 and 18.79, respectively. If we choose a different
training set instead, then we will obtain somewhat different errors on the
validation set.

```{r}
set.seed(2)
train <- sample(392, 196)

lm.fit <- lm(mpg ~ horsepower, subset = train)
mean((mpg - predict(lm.fit, Auto))[-train]^2)

lm.fit2 <- lm(mpg ~ poly(horsepower, 2), data = Auto, subset = train)
mean((mpg - predict(lm.fit2, Auto))[-train]^2)

lm.fit3 <- lm(mpg ~ poly(horsepower, 3), data = Auto, subset = train)
mean((mpg - predict(lm.fit3, Auto))[-train]^2)
```
Using this split of the observations into a training set and a validation
set, we find that the validation set error rates for the models with linear,
quadratic, and cubic terms are 25.73, 20.43, and 20.39, respectively.

These results are consistent with our previous findings: a model that
predicts mpg using a quadratic function of horsepower performs better than
a model that involves only a linear function of horsepower, and there is
little evidence in favor of a model that uses a cubic function of horsepower.
---

## 5.3.2 Leave-One-Out Cross-Validation

The LOOCV estimate can be automatically computed for any generalized
linear model using the glm() and cv.glm() functions. In the lab for Chap- 
ter 4, we used the glm() function to perform logistic regression by passing
in the family = "binomial" argument. But if we use glm() to fit a model
without passing in the family argument, then it performs linear regression,
just like the lm() function. So for instance,

```{r}
glm.fit <- glm(mpg ~ horsepower, data = Auto)
coef(glm.fit)

lm.fit <- lm(mpg ~ horsepower, data = Auto)
coef(lm.fit)
```

yield identical linear regression models. In this lab, we will perform linear
regression using the glm() function rather than the lm() function because
the former can be used together with cv.glm(). The cv.glm() function is
part of the boot library.

```{r}
cv.err <- cv.glm(Auto, glm.fit)
cv.err$delta
```

Thecv.glm() function produces a list with several components.The two
numbers in the delta vector contain the cross-validation results. In this
case the numbers are identical (up to two decimal places) and correspond
to the LOOCV statistic given in (5.1).Below, we discuss a situation in
which the two numbers differ. Our cross-validation estimate for the test
error i sapproximately 24.23.

We can repeat this procedure for increasingly complex polynomial fits.
To automate the process, we use the function to initiate a for loop for()
for loop which iteratively fits polynomial regressions for polynomials of orderi=1
to i=10, computes the associated cross-validation error, and stores it in
the ith element of the vector cv.error. We begin by initializing the vector.
```{r}
cv.error <- rep(0, 10)
for (i in 1:10) {
  glm.fit <- glm(mpg ~ poly(horsepower, i), data = Auto)
  cv.error[i] <- cv.glm(Auto, glm.fit)$delta[1]
}
cv.error
```
As in Figure5.4, we see a sharp drop in the estimated test MSE between
the linear and quadratic fits, but then no clear improvement from using
higher-orderpolynomials.

---

## 5.3.3 k-Fold Cross-Validation

The cv.glm() function can also be used to implement k-fold CV. Below we
use k=10, a common choice for k, on the Auto dataset. We once again set
a random seed and initialize a vector in which we will store the CV errors
corresponding to the polynomial fits of orders one to ten.

```{r}
set.seed(17)
cv.error.10 <- rep(0, 10)
for (i in 1:10) {
  glm.fit <- glm(mpg ~ poly(horsepower, i), data = Auto)
  cv.error.10[i] <- cv.glm(Auto, glm.fit, K = 10)$delta[1]
}
cv.error.10
```

The cv.glm() function can also be used to implement k-fold CV.Below we
use k=10, a common choice fork, on the Auto dataset.We once again set
a random seed and initialize a vector in which we will store the CV errors
corresponding to the polynomial fits of orders one to ten.

We saw in Section 5.3.2 that the two numbers associated with delta are
essentially the same when LOOCV is performed. When we instead perform
k-fold CV, then the two numbers associated with delta differ slightly. The
first is the standard k-fold CV estimate, as in (5.3). The second is a bias
corrected version. On this data set, the two estimates are very similar to
each other.