---
title: "ISLR2 Lab 3.6.3–3.6.6: Linear Regression"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
install.packages("car")
install.packages("ISLR2")
library(ISLR2)
library(car)
```

## 3.6.3 Multiple Linear Regression

In order to fit a multiple linear regression model using least squares, we
again use the lm() function. The syntax lm(y ∼ x1 + x2 + x3) is used to
fit a model with three predictors, x1, x2, and x3. The summary() function
now outputs the regression coefficients for all the predictors.

```{r}
lm.fit <- lm(medv ~ lstat + age, data = Boston)
summary(lm.fit)
```

Use `.` shorthand to rThe Boston data set contains 12 variables, and so it would be cumbersome
to have to type all of these in order to perform a regression using all of the
predictors. Instead, we can use the following short-hand:egress `medv` onto all predictors
 in the dataset.

```{r}
lm.fit <- lm(medv ~ ., data = Boston)
summary(lm.fit)
```
We can access the individual components of a summary object by name
(type ?summary.lm to see what is available). Hence summary(lm.fit)$r.sq
gives us the R2, and summary(lm.fit)$sigma gives us the RSE. The vif()
function, part of the car package, can be used to compute variance inflation
factors. Most VIF’s are low to moderate for this data. The car package is
not part of the base R installation so it must be downloaded the first time
you use it via the install.packages() function in R.

```{r}
summary(lm.fit)$r.sq
summary(lm.fit)$sigma
vif(lm.fit)
```

What if we would like to perform a regression using all of the variables but
one? For example, in the above regression output, age has a high p-value.
So we may wish to run a regression excluding this predictor. The following
syntax results in a regression using all predictors except age.

```{r}
lm.fit1 <- lm(medv ~ . - age, data = Boston)
summary(lm.fit1)
```

Alternatively, the update() function can be used.

```{r}
lm.fit1 <- update(lm.fit, ~ . - age)
```

---

## 3.6.4 Interaction Terms

It is easy to include interaction terms in a linear model using the lm() function. The syntax lstat:age tells R to include an interaction term between
lstat and age. The syntax lstat * age simultaneously includes lstat, age,
and the interaction term lstat×age as predictors; it is a shorthand for
lstat + age + lstat:age

```{r}
summary(lm(medv ~ lstat * age, data = Boston))
```

---

## 3.6.5 Non-linear Transformations of the Predictors

The lm() function can also accommodate non-linear transformations of the
predictors. For instance, given a predictor X, we can create a predictor X2
using I(X^2). The function I() is needed since the ^ has a special meaning I() in a formula object; wrapping as we do allows the standard usage in R,
which is to raise X to the power 2. We now perform a regression of medv
onto lstat and lstat^2.

```{r}
lm.fit2 <- lm(medv ~ lstat + I(lstat^2))
summary(lm.fit2)
```

The near-zero p-value associated with the quadratic term suggests that
it leads to an improved model. We use the anova() function to further 
quantify the extent to which the quadratic fit is superior to the linear fit.

```{r}
lm.fit <- lm(medv ~ lstat)
anova(lm.fit, lm.fit2)
```

Here Model 1 represents the linear submodel containing only one predictor,
lstat, while Model 2 corresponds to the larger quadratic model that has two predictors, lstat and lstat2. The anova() function performs a hypothesis
test comparing the two models. The null hypothesis is that the two models
fit the data equally well, and the alternative hypothesis is that the full
model is superior. Here the F-statistic is 135 and the associated p-value is
virtually zero. This provides very clear evidence that the model containing
the predictors lstat and lstat2 is far superior to the model that only
contains the predictor lstat. This is not surprising, since earlier we saw
evidence for non-linearity in the relationship between medv and lstat. If we
type
```{r}
par(mfrow = c(2, 2))
plot(lm.fit2)
```

then we see that when the lstat2 term is included in the model, there is
little discernible pattern in the residuals. In order to create a cubic fit, we can include a predictor of the form
I(X^3). However, this approach can start to get cumbersome for higherorder polynomials. A better approach involves using the poly() function
to create the polynomial within lm(). For example, the following command
produces a fifth-order polynomial fit:
```{r}
lm.fit5 <- lm(medv ~ poly(lstat, 5))
summary(lm.fit5)
```

This suggests that including additional polynomial terms, up to fifth order,
leads to an improvement in the model fit! However, further investigation
of the data reveals that no polynomial terms beyond fifth order have significant p-values in a regression fit.

By default, the poly() function orthogonalizes the predictors: this means
that the features output by this function are not simply a sequence of
powers of the argument. However, a linear model applied to the output
of the poly() function will have the same fitted values as a linear model
applied to the raw polynomials (although the coefficient estimates, standard
errors, and p-values will differ). In order to obtain the raw polynomials from
the poly() function, the argument raw = TRUE must be used.
Of course, we are in no way restricted to using polynomial transformations of the predictors. 
Here we try a log transformation.

```{r}
summary(lm(medv ~ log(rm), data = Boston))
```

---

## 3.6.6 Qualitative Predictors

We will now examine the Carseats data, which is part of the ISLR2 library.
We will attempt to predict Sales (child car seat sales) in 400 locations
based on a number of predictors.

```{r}
head(Carseats)
```

The Carseats data includes qualitative predictors such as Shelveloc, an
indicator of the quality of the shelving location—that is, the space within
a store in which the car seat is displayed—at each location. The predictor Shelveloc takes on three possible values: Bad, Medium, and Good.
Given a qualitative variable such as Shelveloc, R generates dummy variables
automatically. Below we fit a multiple regression model that includes some
interaction terms.

```{r}
lm.fit <- lm(Sales ~ . + Income:Advertising + Price:Age, data = Carseats)
summary(lm.fit)
```

The contrasts() function returns the coding that R uses for the dummy variables

```{r}
attach(Carseats)
contrasts(ShelveLoc)
```
Use contrasts to learn about other contrasts, and how to set them.

R has created a ShelveLocGood dummy variable that takes on a value of
1 if the shelving location is good, and 0 otherwise. It has also created a
ShelveLocMedium dummy variable that equals 1 if the shelving location is
medium, and 0 otherwise. A bad shelving location corresponds to a zero
for each of the two dummy variables. The fact that the coefficient for
ShelveLocGood in the regression output is positive indicates that a good
shelving location is associated with high sales (relative to a bad location).
And ShelveLocMedium has a smaller positive coefficient, indicating that a
medium shelving location is associated with higher sales than a bad shelving location 
but lower sales than a good shelving location