---
title: "Predicting Daily Citibike Trips"
author: "Your Name"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, message=FALSE}
library(tidyverse)
library(modelr)
```

## Load data

```{r}
trips_per_day <- read_tsv("trips_per_day.tsv")

head(trips_per_day)
```

## Train/test split

Hold out 10% of the data as a final test set. Use the remaining 90% for
training and validation.

```{r}
set.seed(42) # 42 is the answer to the ultimate question of life

# isolate 10% of test set
total_days <- nrow(trips_per_day)
frac_test <- 0.10
num_test <- floor(total_days * frac_test)

# randomly sample rows for the final test set 
test_ndx <- sample(1:total_days, num_test, replace = FALSE)

# used to fit the model
# cutout the final test holdout; lock it in box until the end 
trips_per_day_test <- trips_per_day[test_ndx, ]

# used to evaluate the fit
# remaining 90% becomes the "development" pool
trips_dev_pool <- trips_per_day[-test_ndx, ]

# sanity check: do the counts add up to total days? returns a bool
nrow(trips_per_day_test)
nrow(trips_dev_pool)
nrow(trips_per_day_test) + nrow(trips_dev_pool) == total_days

cat("Dev Pool, train & validate (90%):", nrow(trips_dev_pool), "days\n") 
cat("Final Holdout Test Set (10%):", nrow(trips_per_day_test), "days\n") 
```

## Baseline model: temperature only

Start with the polynomial model from the complexity control notebook.

```{r}
# TODO: fit a linear model using poly(tmin, degree) on training data
# TODO: try different polynomial degrees and evaluate RMSE on validation data

# "k-fold cross-validation if you'd like a more robust estimate"
# letting k = 10 so trains/validates on 9 and tests on 1
# is that too much? how do i even know ... 

set.seed(42)
num_folds <- 10
num_dev_days <- nrow(trips_dev_pool)

trips_dev_pool_folds <- trips_dev_pool %>%
    slice_sample(prop = 1) %>%
    mutate(fold = (row_number() %% num_folds) + 1)

# fit a model for each polynomial degree
K <- 1:8
avg_validate_err <- c()
se_validate_err <- c()

for (k in K) {

  # do 10-fold cross-validation within each value of k
  validate_err <- c()
  for (f in 1:num_folds) {
    # fit on the training data
    trips_per_day_train <- filter(trips_dev_pool_folds, fold != f)
    model <- lm(num_trips ~ poly(tmin, k, raw = TRUE), data = trips_per_day_train)


    # evaluate on the validation data
    trips_per_day_validate <- filter(trips_dev_pool_folds, fold == f)
    validate_err[f] <- sqrt(mean((predict(model, trips_per_day_validate) - trips_per_day_validate$num_trips)^2))
  }

  # compute the average validation error across folds
  # and the standard error on this estimate
  avg_validate_err[k] <- mean(validate_err)
  se_validate_err[k] <- sd(validate_err) / sqrt(num_folds)
}

baseline_results <- tibble(
  degree = K,
  avg_validate_rmse = avg_validate_err,
  standard_error = se_validate_err
)

print(baseline_results)

# plot the validate error, highlighting the value of k with the lowest average error
plot_data <- data.frame(K, avg_validate_err, se_validate_err)
ggplot(plot_data, aes(x=K, y=avg_validate_err)) +
  geom_pointrange(aes(ymin=avg_validate_err - se_validate_err,
                      ymax=avg_validate_err + se_validate_err,
                      color=avg_validate_err == min(avg_validate_err))) +
  geom_line(color = "red") +
  scale_x_continuous(breaks=1:10) +
  theme(legend.position="none") +
  xlab('Polynomial Degree') +
  ylab('RMSE on validation data')

# I'm going with degree 4 instead of degree 5
# 4      4             5569.           198.
# 5      5             5561.           201.
# hoping it's less likely to overfit 

```

## Feature engineering

Add features that might help: day of week, month, precipitation, etc.
Remember: only use information available *before* the day in question.

```{r}

```

## Model comparison

Try different models and compare their validation performance.

```{r}

```

## Final model

Pick your best model and plot the results.

```{r}

```

## Save model

```{r}

```

## Test set evaluation

**Do this only once, at the very end!**

```{r}
# TODO: load your saved model and predict on trips_test
# TODO: compute RMSE on test set
# TODO: how do you think this model will perform on future data?
```
