---
title: "Predicting Daily Citibike Trips"
author: "Your Name"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, message=FALSE}
library(tidyverse)
library(modelr)
```

## Load data

```{r}
trips_per_day <- read_tsv("trips_per_day.tsv")

head(trips_per_day)
```

## Train/test split

Hold out 10% of the data as a final test set. Use the remaining 90% for
training and validation.

```{r}
set.seed(42) # 42 is the answer to the ultimate question of life

# isolate 10% of test set
total_days <- nrow(trips_per_day)
frac_test <- 0.10
num_test <- floor(total_days * frac_test)

# randomly sample rows for the final test set 
test_ndx <- sample(1:total_days, num_test, replace = FALSE)

# used to fit the model
# cutout the final test holdout; lock it in box until the end 
trips_per_day_test <- trips_per_day[test_ndx, ]

# used to evaluate the fit
# remaining 90% becomes the "development" pool
trips_dev_pool <- trips_per_day[-test_ndx, ]

# sanity check: do the counts add up to total days? returns a bool
nrow(trips_per_day_test)
nrow(trips_dev_pool)
nrow(trips_per_day_test) + nrow(trips_dev_pool) == total_days

cat("Dev Pool, train & validate (90%):", nrow(trips_dev_pool), "days\n") 
cat("Final Holdout Test Set (10%):", nrow(trips_per_day_test), "days\n") 
```

## Baseline model: temperature only

Start with the polynomial model from the complexity control notebook.

```{r}
# TODO: fit a linear model using poly(tmin, degree) on training data
# TODO: try different polynomial degrees and evaluate RMSE on validation data

# "k-fold cross-validation if you'd like a more robust estimate"
# letting k = 10 so trains/validates on 9 and tests on 1
# is that too much? how do i even know ... 
```

## Feature engineering

Add features that might help: day of week, month, precipitation, etc.
Remember: only use information available *before* the day in question.

```{r}

```

## Model comparison

Try different models and compare their validation performance.

```{r}

```

## Final model

Pick your best model and plot the results.

```{r}

```

## Save model

```{r}

```

## Test set evaluation

**Do this only once, at the very end!**

```{r}
# TODO: load your saved model and predict on trips_test
# TODO: compute RMSE on test set
# TODO: how do you think this model will perform on future data?
```
