---
title: "Predicting Daily Citibike Trips"
author: "Your Name"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, message=FALSE}
library(tidyverse)
library(modelr)
```

## Load data

```{r}
trips_per_day <- read_tsv("trips_per_day.tsv")

head(trips_per_day)
```

## Train/test split

Hold out 10% of the data as a final test set. Use the remaining 90% for
training and validation.

```{r}
set.seed(42) # 42 is the answer to the ultimate question of life

# isolate 10% of test set
total_days <- nrow(trips_per_day)
frac_test <- 0.10
num_test <- floor(total_days * frac_test)

# randomly sample rows for the final test set 
test_ndx <- sample(1:total_days, num_test, replace = FALSE)

# used to fit the model
# cutout the final test holdout; lock it in box until the end 
trips_per_day_test <- trips_per_day[test_ndx, ]

# used to evaluate the fit
# remaining 90% becomes the "development" pool
trips_dev_pool <- trips_per_day[-test_ndx, ]

# sanity check: do the counts add up to total days? returns a bool
nrow(trips_per_day_test)
nrow(trips_dev_pool)
nrow(trips_per_day_test) + nrow(trips_dev_pool) == total_days

cat("Dev Pool, train & validate (90%):", nrow(trips_dev_pool), "days\n") 
cat("Final Holdout Test Set (10%):", nrow(trips_per_day_test), "days\n") 
```

## Baseline model: temperature only

Start with the polynomial model from the complexity control notebook.

```{r}
# TODO: fit a linear model using poly(tmin, degree) on training data
# TODO: try different polynomial degrees and evaluate RMSE on validation data

# "k-fold cross-validation if you'd like a more robust estimate"
# letting k = 10 so trains/validates on 9 and tests on 1
# is that too much? how do i even know ... 

set.seed(42)
num_folds <- 10
num_dev_days <- nrow(trips_dev_pool)

trips_dev_pool_folds <- trips_dev_pool %>%
    slice_sample(prop = 1) %>%
    mutate(fold = (row_number() %% num_folds) + 1)

# fit a model for each polynomial degree
K <- 1:8
avg_validate_err <- c()
se_validate_err <- c()

for (k in K) {

  # do 10-fold cross-validation within each value of k
  validate_err <- c()
  for (f in 1:num_folds) {
    # fit on the training data
    trips_per_day_train <- filter(trips_dev_pool_folds, fold != f)
    model <- lm(num_trips ~ poly(tmin, k, raw = TRUE), data = trips_per_day_train)


    # evaluate on the validation data
    trips_per_day_validate <- filter(trips_dev_pool_folds, fold == f)
    validate_err[f] <- sqrt(mean((predict(model, trips_per_day_validate) - trips_per_day_validate$num_trips)^2))
  }

  # compute the average validation error across folds
  # and the standard error on this estimate
  avg_validate_err[k] <- mean(validate_err)
  se_validate_err[k] <- sd(validate_err) / sqrt(num_folds)
}

baseline_results <- tibble(
  degree = K,
  avg_validate_rmse = avg_validate_err,
  standard_error = se_validate_err
)

print(baseline_results)

# plot the validate error, highlighting the value of k with the lowest average error
plot_data <- data.frame(K, avg_validate_err, se_validate_err)
ggplot(plot_data, aes(x=K, y=avg_validate_err)) +
  geom_pointrange(aes(ymin=avg_validate_err - se_validate_err,
                      ymax=avg_validate_err + se_validate_err,
                      color=avg_validate_err == min(avg_validate_err))) +
  geom_line(color = "red") +
  scale_x_continuous(breaks=1:8) +
  theme(legend.position="none") +
  xlab('Polynomial Degree') +
  ylab('RMSE on validation data')

# I'm going with degree 4 instead of degree 5
# 4      4             5569.           198.
# 5      5             5561.           201.
# hoping it's less likely to overfit 

```

## Feature engineering

Add features that might help: day of week, month, precipitation, etc.
Remember: only use information available *before* the day in question.

```{r}

# i just created a new file called US_holidays.csv and copy and pasted
# first row added col names; date, holiday

# Update: I can't get it to work so I'm explicitly adding federal holidays 
holiday_dates <- ymd(c(
  "2014-01-01", # New Year's Day
  "2014-01-20", # Martin Luther King Jr. Day
  "2014-02-17", # Presidents' Day
  "2014-05-26", # Memorial Day
  "2014-07-04", # Independence Day
  "2014-09-01", # Labor Day
  "2014-10-13", # Columbus Day
  "2014-11-11", # Veterans Day
  "2014-11-27", # Thanksgiving
  "2014-12-25"  # Christmas Day
))
# I put 10 holidays but this output
# A tibble: 2 × 2
# is_holiday     n
#      <dbl> <int>
#1          0   321
#2          1     8
# why is it only catching 8 ??

trips_dev_pool_creative <- trips_dev_pool %>%
  arrange(date) %>%
  mutate(date = ymd(date)) %>% # convert to date
  
  # Temperature Metrics
  mutate(
    temp_range = tmax - tmin
  ) %>%

  # Add-on 1: Time & Commute
  mutate(
    day_of_week = wday(date, label = TRUE),
    is_weekend = if_else(day_of_week %in% c("Sat", "Sun"), 1, 0),
    month_of_year = month(date, label = TRUE)
  ) %>%

  # Add-on 2: Environmental Friction
  mutate(
    is_raining = if_else(prcp > 0, 1, 0),
    is_snowing = if_else(snow > 0, 1, 0),
    has_snow_depth = if_else(snwd > 0, 1, 0)
  ) %>%
  
  # Add-on 3: Holiday Effects
  mutate(
    is_holiday = if_else(date %in% holiday_dates, 1, 0)
  )

# Check whats up
trips_dev_pool_creative %>% count(is_holiday)
glimpse(trips_dev_pool_creative)
```

```{r}
set.seed(42)
num_folds <- 10

# Scramble dataset and assign folds
trips_dev_pool_folds <- trips_dev_pool_creative %>%
    slice_sample(prop = 1) %>%
    mutate(fold = (row_number() %% num_folds) + 1)

# init vectors
err_baseline <- c()
err_time <- c()
err_env_friction <- c()
err_holiday <- c()
err_all_features <- c()

# Run 10-fold
for (f in 1:num_folds) {
    # Separate into training and validation folds
    train <- filter(trips_dev_pool_folds, fold != f)
    validate <- filter(trips_dev_pool_folds, fold == f)

    # Experiment 0: Baseline (Temp Only)
    m_baseline <- lm(num_trips ~ poly(tmin, 4, raw = TRUE), data = train)
    err_baseline[f] <- sqrt(mean((predict(m_baseline, validate) - validate$num_trips)^2))

    # Experiment 1: Temp + Time (Commute Patterns)
    m_time <- lm(num_trips ~ poly(tmin, 4, raw = TRUE) * is_weekend + day_of_week + month_of_year, data = train)
    err_time[f] <- sqrt(mean((predict(m_time, validate) - validate$num_trips)^2))
 
    # Experiment 2: Temperature + Environmental Friction
    m_env_friction <- lm(num_trips ~ poly(tmin, 4, raw = TRUE) + is_raining + is_snowing + has_snow_depth, data = train)
    err_env_friction[f] <- sqrt(mean((predict(m_env_friction, validate) - validate$num_trips)^2))
 
    # Experiment 3: Temperature + Holiday Effects
    m_holiday <- lm(num_trips ~ poly(tmin, 4, raw = TRUE) + is_holiday, data = train)
    err_holiday[f] <- sqrt(mean((predict(m_holiday, validate) - validate$num_trips)^2))
 
    # Experiment 4: Combined Model (All Features)
    m_all_features <- lm(num_trips ~ poly(tmin, 4, raw = TRUE) * is_weekend +
                                    day_of_week + month_of_year +
                                    is_raining + is_snowing + has_snow_depth +
                                    is_holiday, data = train)
    err_all_features[f] = sqrt(mean((predict(m_all_features, validate) - validate$num_trips)^2))
}
```

## Model comparison

Try different models and compare their validation performance.

```{r}
experiment_results <- tibble(
  experiment = c(
    "baseline_temp_only", 
    "addon_1_temp_time", 
    "addon_2_temp_envfrict", 
    "addon_3_temp_holidays", 
    "all_addons"
  ),
  avg_validate_RMSE = c(
    mean(err_baseline), 
    mean(err_time), 
    mean(err_env_friction), 
    mean(err_holiday), 
    mean(err_all_features)
  ),
  standard_error = c(
    sd(err_baseline) / sqrt(num_folds), 
    sd(err_time) / sqrt(num_folds), 
    sd(err_env_friction) / sqrt(num_folds), 
    sd(err_holiday) / sqrt(num_folds), 
    sd(err_all_features) / sqrt(num_folds)
  )
)

# arrange by lowest error aka best performance
experiment_results <- experiment_results %>% 
  arrange(avg_validate_RMSE)

# view final comparison table
print(experiment_results)

# concerns: will this overfit on unseen data?
# environmental friction is the most stable add on; lowest standard error
# all add ons combined is the most "powerful"; lowest avg RMSE
```

## Final model

Pick your best model and plot the results.
Going with all_addons
avg_validate RMSE of 3963
standard_error of 195

all_addons is the Combined Model (All Features)
train final model on trips_dev_pool_creative

```{r}
final_model <- lm(num_trips ~ poly(tmin, 4, raw = TRUE) * is_weekend +
                              day_of_week + month_of_year +
                              is_raining + is_snowing + has_snow_depth +
                              is_holiday, 
                  data = trips_dev_pool_creative)
                  
trips_dev_pool_plots <- 
    add_predictions(trips_dev_pool_creative, final_model, var = "predicted_trips", type = NULL)

ggplot(trips_dev_pool_plots, aes(x = predicted_trips, y = num_trips)) +
  geom_point(alpha = 0.6, color = "purple") +
  geom_abline(intercept = 0, slope = 1, color = "orange", linetype = "dashed", size = 1) +
  theme_minimal() +
  labs(
    title = "Actual vs. Predicted Daily Trips",
    subtitle = "Dashed orange line represents a perfect 1:1 prediction",
    x = "Predicted Number of Trips",
    y = "Actual Number of Trips"
  )
```

## Save model

```{r}

```

## Test set evaluation

**Do this only once, at the very end!**

```{r}
# TODO: load your saved model and predict on trips_test
# TODO: compute RMSE on test set
# TODO: how do you think this model will perform on future data?
```